---
title: "R Descriptive Statistics"
author: "Cuneyt Akcora"
date: "January 18, 2019"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(ggplot2)
```

# Contents
1. [Mean and Variance](#mean-and-variance)
2. [Probability Plot](#probability-plot)
3. [Box Plot](#box-plot) 

## Mean and Variance

Caution: The built-in var() function of R uses n-1 in the variance denominatior

```{r variance}
# Let's create a population from a normal distribution. 
set.seed(123)
population<-rnorm(10000,mean=25,sd=3)
sample<-sample(population, 10, replace = FALSE)
sample
mean(population)
mean(sample)
varPopulation= (sum((population-mean(population))^2))/length(population)
varPopulation
#Let's see how sample size changes the results
for(sampleSize in c(10,20,30,90)){
  sample<-sample(population, sampleSize, replace = FALSE)
  varN= (sum((sample-mean(sample))^2))/(sampleSize-1)
  message(sampleSize," data points sampled,  mean is ",mean(sample)," var is ",varN)
}

#Impact of n-1 on the variance estimate
sampleSize=10
#scenario 1: no correction
variances1<-c()
for(samplingRate in 1:1000) {
  sample<-sample(population, sampleSize, replace = FALSE)
  varN= (sum((sample-mean(sample))^2))/sampleSize
  variances1<-cbind(variances1,varN)
}
#scenario 2: n-1 correction
variances2<-c()
for(samplingRate in 1:1000) {
  sample<-sample(population, sampleSize, replace = FALSE)
  varN= (sum((sample-mean(sample))^2))/(sampleSize-1)
  variances2<-cbind(variances2,varN)
}
#scenario 3: n-2 correction
variances3<-c()
for(samplingRate in 1:1000) {
  sample<-sample(population, sampleSize, replace = FALSE)
  varN= (sum((sample-mean(sample))^2))/(sampleSize-2)
  variances3<-cbind(variances3,varN)
}
message("Mean estimated variances are n:",mean(variances1), " n-1:", mean(variances2)," n-2:",mean(variances3))
```

Using n always underestimates, whereas n-2 overestimates. The closest estimation is from the n-1 approach (scenario 2).
 

## Probability Plot

In the almost perfect normally distributed sample, the QQ plot is a straight line.
 
```{r probability}
norm <- rnorm(100000)
norm<-cbind.data.frame(norm,1:length(norm))
colnames(norm)<-c("val","index")

plot2<-ggplot(norm, aes(x=val)) +geom_histogram(bins=50,color="black", fill="lightblue")+
      labs(x="Value", y = "Count")
print(plot2)
probPlot<-ggplot(norm, aes(sample=val))+stat_qq()
print(probPlot)
require(gridExtra)
grid.arrange(probPlot, plot2+coord_flip(), ncol=2)
```

Adding a few too large data points

```{r large}
norm<-rbind(norm,c(5,length((norm))))
norm<-rbind(norm,c(6,length((norm)))) 
norm<-rbind(norm,c(7,length((norm))))
probPlot<-ggplot(norm, aes(sample=val))+stat_qq()
print(probPlot)
```

Adding a few too small data points.

```{r small}
norm<-rbind(norm,c(-5,length((norm))))
norm<-rbind(norm,c(-6,length((norm)))) 
norm<-rbind(norm,c(-7,length((norm))))
probPlot<-ggplot(norm, aes(sample=val))+stat_qq()
print(probPlot)
```

## Box Plot

```{r box}
norm <- rnorm(100000,mean=0,sd=1)
norm<-cbind.data.frame(norm,1:length(norm))
colnames(norm)<-c("norm1","index")
norm2<-rnorm(nrow(norm),mean=0,sd=5)
norm<-cbind(norm,norm2)
plot3<-ggplot(norm,aes(y=norm1))+geom_boxplot(fill = "white", colour = "#3366FF")
plot4<-ggplot(norm,aes(y=norm2))+geom_boxplot(fill = "white", colour = "#3366FF")
grid.arrange(plot4, plot3, ncol=2)
 
```

 